```{r}
library(readxl)
library(pscl)
library(MASS)
library(lmtest)
library(epitools)
library(risks)
library(coenocliner)
library(ggplot2)
library(vcdExtra)
library(DHARMa)
library(countreg)
library(writexl)
library(foreign)
library(sandwich)
library(glmmADMB)
```


# Schistosomiasis count data (STH)
```{r}
count_data <- read_excel("Dataframe.xlsx")
```
# Column names
```{r}
colnames(count_data)
```
# Mean of schistosomiasis count data (with zeros and without zeros)
```{r}
print(paste("The mean of the counts with zeros is ", mean(count_data$Schistosomiasis)))
print(paste("The mean of the counts without zeros is ", mean(count_data$Schistosomiasis>0)))
```
```{r}
five_predictors = 'Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation'
ten_predictors = 'Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole'
```

# USING ALL SAMPLE (ALL INTENSITY) DATA

## The Poisson model
```{r}
poisson.model_all5 <- glm(five_predictors, family = 'poisson', data = count_data)
summary(poisson.model_all5)
```

```{r}
poisson.model_all10 <- glm(ten_predictors, family = 'poisson', data = count_data)
summary(poisson.model_all10)
```
## The Negative Binomial model
```{r}
negbin.model_all5 <- glm(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation, family = negative.binomial(1), data = count_data)
summary(negbin.model_all5)
```

```{r}
negbin.model_all10 <- glm(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, family = negative.binomial(1), data = count_data)
summary(negbin.model_all10)
```

## The zero inflated Poisson model
```{r}
zip.model_all5 = zeroinfl(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation, data = count_data, dist = "poisson")
summary(zip.model_all5)
```

```{r}
zip.model_all10 = zeroinfl(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = count_data, dist = "poisson")
summary(zip.model_all10)
```


## The zero negative binomial model
```{r}
zinb.model_all5 = zeroinfl(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation, data = count_data, dist = "negbin")
summary(zinb.model_all5)
```

```{r}
zinb.model_all10 = zeroinfl(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = count_data, dist = "negbin")
summary(zinb.model_all10)
```


## Hurdle Poisson model (HP)
```{r}
hurdlep.model_all5 = hurdle(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation, data = count_data, dist = "poisson")
summary(hurdlep.model_all5)
```

```{r}
hurdlep.model_all10 = hurdle(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = count_data, dist = "poisson")
summary(hurdlep.model_all10)
```

## Hurdle negative binomial model
```{r}
hurdlenb.model_all5 = hurdle(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation, data = count_data, dist = "negbin")
summary(hurdlenb.model_all5)
```


```{r}
hurdlenb.model_all10 = hurdle(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = count_data, dist = "negbin")
summary(hurdlenb.model_all10)
```






# USING LOW INTENSITY DATA

```{r}
low_insensity_data = count_data[count_data$Schistosomiasis<100,]
```

## Poisson model

```{r}
poisson.model_low5 <- glm(five_predictors, family = 'poisson', data = low_insensity_data)
summary(poisson.model_low5)
```
```{r}
poisson.model_low10 <- glm(ten_predictors, family = 'poisson', data = low_insensity_data)
summary(poisson.model_low10)
```

## Negative binomial model
```{r}
negbin.model_low5 <- glm(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation, family = negative.binomial(1), data = low_insensity_data)
summary(negbin.model_low5)
```


```{r}
negbin.model_low10 <- glm(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, family = negative.binomial(1), data = low_insensity_data)
summary(negbin.model_low10)
```

## Zero-inflated Poisson model
```{r}
zip.model_low5 = zeroinfl(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation, data = low_insensity_data, dist = "poisson")
summary(zip.model_low5)
```

```{r}
zip.model_low10 = zeroinfl(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = low_insensity_data, dist = "poisson")
summary(zip.model_low10)
```

## The zero-inflated negative binomial model
```{r}
zinb.model_low5 = zeroinfl(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation, data = low_insensity_data, dist = "negbin")
summary(zinb.model_low5)
```



```{r}
zinb.model_low10 = zeroinfl(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = low_insensity_data, dist = "negbin")
summary(zinb.model_low10)
```

## Hurdle Poisson model
```{r}
hurdlep.model_low5 = hurdle(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation, data = low_insensity_data, dist = "poisson")
summary(hurdlep.model_low5)
```

```{r}
hurdlep.model_low10 = hurdle(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = low_insensity_data, dist = "poisson")
summary(hurdlep.model_low10)
```

## Hurdle negative binomial model
```{r}
hurdlenb.model_low5 = hurdle(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation, data = low_insensity_data, dist = "negbin")
summary(hurdlenb.model_low5)
```


```{r}
hurdlenb.model_low10 = hurdle(Schistosomiasis ~ Age + Sex + Class + Community + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = low_insensity_data, dist = "negbin")
summary(hurdlenb.model_low10)
```







































# ANALYSIS

## AIC and BIC values for using all intensity samples
```{r}
AIC_all5 = AIC(poisson.model_all5, negbin.model_all5, zip.model_all5, zinb.model_all5, hurdlep.model_all5, hurdlenb.model_all5)
write_xlsx(AIC_all5,"C:\\Users\\User\\Desktop\\Model AIC and BIC values\\AIC_all5.xlsx")

AIC_all10 = AIC(poisson.model_all10, negbin.model_all10, zip.model_all10, zinb.model_all10, hurdlep.model_all10, hurdlenb.model_all10)
write_xlsx(AIC_all10,"C:\\Users\\User\\Desktop\\Model AIC and BIC values\\AIC_all10.xlsx")
```

```{r}
BIC_all5 = BIC(poisson.model_all5, negbin.model_all5, zip.model_all5, zinb.model_all5, hurdlep.model_all5, hurdlenb.model_all5)
#write_xlsx(BIC_all5,"C:\\Users\\User\\Desktop\\Model AIC and BIC values\\BIC_all5.xlsx")


BIC_all10 = BIC(poisson.model_all10, negbin.model_all10, zip.model_all10, zinb.model_all10, hurdlep.model_all10, hurdlenb.model_all10)
#write_xlsx(BIC_all10,"C:\\Users\\User\\Desktop\\Model AIC and BIC values\\BIC_all10.xlsx")
```

## AIC and BIC values for using low intensity samples
```{r}
AIC_low5 = AIC(poisson.model_low5, negbin.model_low5, zip.model_low5, zinb.model_low5, hurdlep.model_low5, hurdlenb.model_low5)
write_xlsx(AIC_low5,"C:\\Users\\User\\Desktop\\Model AIC and BIC values\\AIC_low5.xlsx")

AIC_low10 = AIC(poisson.model_low10, negbin.model_low10, zip.model_low10, zinb.model_low10, hurdlep.model_low10, hurdlenb.model_low10)
write_xlsx(AIC_low10,"C:\\Users\\User\\Desktop\\Model AIC and BIC values\\AIC_low10.xlsx")
```

```{r}
BIC_low5 = BIC(poisson.model_low5, negbin.model_low5, zip.model_low5, zinb.model_low5, hurdlep.model_low5, hurdlenb.model_low5)
#write_xlsx(BIC_low5,"C:\\Users\\User\\Desktop\\Model AIC and BIC values\\BIC_low5.xlsx")

BIC_low10 = BIC(poisson.model_low10, negbin.model_low10, zip.model_low10, zinb.model_low10, hurdlep.model_low10, hurdlenb.model_low10)
#write_xlsx(BIC_low10,"C:\\Users\\User\\Desktop\\Model AIC and BIC values\\BIC_low10.xlsx")
```














```{r}
hurdlenb.model_all10$theta
```



## Getting model's observed and predicted values

```{r}
poisson_results = rootogram(poisson.model_all10, max = 50, style = "standing", plot = FALSE)
negbin_results <- rootogram(negbin.model_all10, max = 50, style = "standing", plot = FALSE)
zip_results = rootogram(zip.model_all10, max = 50, style = "standing", plot = FALSE)
zinb_results = rootogram(zinb.model_all10, max = 50, style = "standing", plot = FALSE)
hurdlePoisson_results = rootogram(hurdlep.model_all10, max = 50, style = "standing", plot = FALSE)
hurdleNB_results = rootogram(hurdlenb.model_all10, max = 50, style = "standing", plot = FALSE)
```



## Saving model results
uncomment to have the output results
```{r}
write_xlsx(poisson_results,"C:\\Users\\User\\Desktop\\Model Results\\poisson_results.xlsx")
write_xlsx(negbin_results,"C:\\Users\\User\\Desktop\\Model Results\\negbin_results.xlsx")
write_xlsx(zip_results,"C:\\Users\\User\\Desktop\\Model Results\\zip_results.xlsx")
write_xlsx(zinb_results,"C:\\Users\\User\\Desktop\\Model Results\\zinb_results.xlsx")
write_xlsx(hurdlePoisson_results,"C:\\Users\\User\\Desktop\\Model Results\\hurdlePoisson_results.xlsx")
write_xlsx(hurdleNB_results,"C:\\Users\\User\\Desktop\\Model Results\\hurdleNB_results.xlsx")
```


## Coefficients of hurdle model's predictors
```{r}
print(coef(hurdlenb.model_all10))
```

## Calculating the odds ratio and 95% confidence interval for each predictor variable from the hurdle NB model's results
```{r}
exp(cbind(Odds_Ratio = coef(hurdlenb.model_all10), confint(hurdlenb.model_all10)))
```

## The likelihood ratio test for nested models 
```{r}
lrtest(poisson.model_all10, poisson.model_all5)
```

```{r}
lrtest(negbin.model_all10, negbin.model_all5)
```
```{r}
lrtest(zip.model_all10, zip.model_all5)
```


```{r}
lrtest(zinb.model_all10, zinb.model_all5)
```

```{r}
lrtest(hurdlep.model_all10, hurdlep.model_all5)
```

```{r}
lrtest(hurdlenb.model_all10,hurdlenb.model_all5)
```

## The vuoung test between ZINB and hurdle NB (ZINB & HNB)
```{r}
vuong(hurdlenb.model_all10, zinb.model_all10)
```





