```{r}
library(readxl)
library(pscl)
library(MASS)
library(lmtest)
library(epitools)
library(risks)
library(coenocliner)
library(ggplot2)
library(vcdExtra)
library(DHARMa)
library(countreg)
library(writexl)
library(foreign)
library(sandwich)
library(glmmADMB)
library(AER)
library(foreign)
library(xtable)
library(epiDisplay)
library(gtsummary)
library(htmlTable)
library(xtable)
options(scipen=999)
```


# loading the Schistosomiasis count data (STH)

```{r}
count_data <- read_excel("hurdle_dataset_new.xlsx")
```

# Column names



```{r}
colnames(count_data)
```

# Mean of schistosomiasis count data (with zeros and without zeros)

```{r}
print(paste("The mean of the counts with zeros is ", mean(count_data$schistosomiasis)))
print(paste("The mean of the counts without zeros is ", mean(count_data$schistosomiasis>0)))
```
The median age

```{r}
median(count_data$Age)
```

# USING ALL SAMPLE (ALL INTENSITY) DATA

Simulating the model with all sample in the data, which is considered as all intensity as it includes low, medium and high intensity counts

The names are assigned to the model in the order {model type}.model_{sample size}{number of predictors}

## The Poisson model

```{r}
poisson.model_all5 <- glm("schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation", family = 'poisson', data = count_data)
summary(poisson.model_all5)
```

```{r}
poisson.model_all10 <- glm("schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole", family = 'poisson', data = count_data)
summary(poisson.model_all10)
```

## The Negative Binomial model

```{r}
negbin.model_all5 <- glm("schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation", family = negative.binomial(1), data = count_data)
summary(negbin.model_all5)
```

```{r}
negbin.model_all10 <- glm("schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole", family = negative.binomial(1), data = count_data)
summary(negbin.model_all10)
```

## The zero inflated Poisson model

```{r}
zip.model_all5 = zeroinfl(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation, data = count_data, dist = "poisson")
summary(zip.model_all5)
```

```{r}
zip.model_all10 = zeroinfl(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = count_data, dist = "poisson")
summary(zip.model_all10)
```

## The zero negative binomial model

```{r}
zinb.model_all5 = zeroinfl(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation, data = count_data, dist = "negbin")
summary(zinb.model_all5)
```

```{r}
zinb.model_all10 = zeroinfl(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = count_data, dist = "negbin")
summary(zinb.model_all10)
```

## Hurdle Poisson model (HP)

```{r}
hurdlep.model_all5 = hurdle(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation, data = count_data, dist = "poisson")
summary(hurdlep.model_all5)
```

```{r}
hurdlep.model_all10 = hurdle(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = count_data, dist = "poisson")
summary(hurdlep.model_all10)
```

## Hurdle negative binomial model

```{r}
hurdlenb.model_all5 = hurdle(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation, data = count_data, dist = "negbin")
summary(hurdlenb.model_all5)
```

```{r}
hurdlenb.model_all10 = hurdle(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = count_data, dist = "negbin")
summary(hurdlenb.model_all10)
```

# USING LOW INTENSITY DATA

```{r}
low_insensity_data = count_data[count_data$schistosomiasis<100,]
```

## Poisson model

```{r}
poisson.model_low5 <- glm(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation, family = 'poisson', data = low_insensity_data)
summary(poisson.model_low5)
```

```{r}
poisson.model_low10 <- glm(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, family = 'poisson', data = low_insensity_data)
summary(poisson.model_low10)
```

## Negative binomial model

```{r}
negbin.model_low5 <- glm(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation, family = negative.binomial(1), data = low_insensity_data)
summary(negbin.model_low5)
```

```{r}
negbin.model_low10 <- glm(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, family = negative.binomial(1), data = low_insensity_data)
summary(negbin.model_low10)
```

## Zero-inflated Poisson model

```{r}
zip.model_low5 = zeroinfl(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation, data = low_insensity_data, dist = "poisson")
summary(zip.model_low5)
```

```{r}
zip.model_low10 = zeroinfl(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = low_insensity_data, dist = "poisson")
summary(zip.model_low10)
```

## The zero-inflated negative binomial model

```{r}
zinb.model_low5 = zeroinfl(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation, data = low_insensity_data, dist = "negbin")
summary(zinb.model_low5)
```

```{r}
zinb.model_low10 = zeroinfl(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = low_insensity_data, dist = "negbin")
summary(zinb.model_low10)
```

## Hurdle Poisson model

```{r}
hurdlep.model_low5 = hurdle(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation, data = low_insensity_data, dist = "poisson")
summary(hurdlep.model_low5)
```

```{r}
hurdlep.model_low10 = hurdle(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = low_insensity_data, dist = "poisson")
summary(hurdlep.model_low10)
```

## Hurdle negative binomial model

```{r}
hurdlenb.model_low5 = hurdle(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation, data = low_insensity_data, dist = "negbin")
summary(hurdlenb.model_low5)
```

```{r}
hurdlenb.model_low10 = hurdle(schistosomiasis ~ Age_group + Sex + Area + EduLevel + Parent_Occupation + Pipe_borne + Tanker_treated + Tanker_Untreated + River_Stream + Well_Borehole, data = low_insensity_data, dist = "negbin")
summary(hurdlenb.model_low10)
```

# ANALYSIS

## AIC and BIC values for using all intensity samples

```{r}
AIC_all5 = AIC(poisson.model_all5, negbin.model_all5, zip.model_all5, zinb.model_all5, hurdlep.model_all5, hurdlenb.model_all5)
write_xlsx(AIC_all5,"aic_bic_values\\AIC_all5.xlsx")

AIC_all10 = AIC(poisson.model_all10, negbin.model_all10, zip.model_all10, zinb.model_all10, hurdlep.model_all10, hurdlenb.model_all10)
write_xlsx(AIC_all10,"aic_bic_values\\AIC_all10.xlsx")
```

```{r}
BIC_all5 = BIC(poisson.model_all5, negbin.model_all5, zip.model_all5, zinb.model_all5, hurdlep.model_all5, hurdlenb.model_all5)
write_xlsx(BIC_all5,"aic_bic_values\\BIC_all5.xlsx")

BIC_all10 = BIC(poisson.model_all10, negbin.model_all10, zip.model_all10, zinb.model_all10, hurdlep.model_all10, hurdlenb.model_all10)
write_xlsx(BIC_all10,"aic_bic_values\\BIC_all10.xlsx")
```

## AIC and BIC values for using low intensity samples

```{r}
AIC_low5 = AIC(poisson.model_low5, negbin.model_low5, zip.model_low5, zinb.model_low5, hurdlep.model_low5, hurdlenb.model_low5)
write_xlsx(AIC_low5,"aic_bic_values\\AIC_low5.xlsx")

AIC_low10 = AIC(poisson.model_low10, negbin.model_low10, zip.model_low10, zinb.model_low10, hurdlep.model_low10, hurdlenb.model_low10)
write_xlsx(AIC_low10,"aic_bic_values\\AIC_low10.xlsx")
```

```{r}
BIC_low5 = BIC(poisson.model_low5, negbin.model_low5, zip.model_low5, zinb.model_low5, hurdlep.model_low5, hurdlenb.model_low5)
write_xlsx(BIC_low5,"aic_bic_values\\BIC_low5.xlsx")

BIC_low10 = BIC(poisson.model_low10, negbin.model_low10, zip.model_low10, zinb.model_low10, hurdlep.model_low10, hurdlenb.model_low10)
write_xlsx(BIC_low10,"aic_bic_values\\BIC_low10.xlsx")
```

```{r}
hurdlenb.model_all10$theta
```

## Getting model's observed and predicted values

```{r}
poisson_results = rootogram(poisson.model_all10, max = 50, style = "standing", plot = FALSE)
negbin_results <- rootogram(negbin.model_all10, max = 50, style = "standing", plot = FALSE)
zip_results = rootogram(zip.model_all10, max = 50, style = "standing", plot = FALSE)
zinb_results = rootogram(zinb.model_all10, max = 50, style = "standing", plot = FALSE)
hurdlePoisson_results = rootogram(hurdlep.model_all10, max = 50, style = "standing", plot = FALSE)
hurdleNB_results = rootogram(hurdlenb.model_all10, max = 50, style = "standing", plot = FALSE)
```

## Saving model results

uncomment to get the output results

```{r}
write_xlsx(poisson_results,"C:\\Users\\User\\Desktop\\paper review\\model results\\poisson_results.xlsx")
write_xlsx(negbin_results,"C:\\Users\\User\\Desktop\\paper review\\model results\\negbin_results.xlsx")
write_xlsx(zip_results,"C:\\Users\\User\\Desktop\\paper review\\model results\\zip_results.xlsx")
write_xlsx(zinb_results,"C:\\Users\\User\\Desktop\\paper review\\model results\\zinb_results.xlsx")
write_xlsx(hurdlePoisson_results,"C:\\Users\\User\\Desktop\\paper review\\model results\\hurdlePoisson_results.xlsx")
write_xlsx(hurdleNB_results,"C:\\Users\\User\\Desktop\\paper review\\model results\\hurdleNB_results.xlsx")
```

## Coefficients of hurdle model's predictors

```{r}
print(coef(hurdlenb.model_all10))
```

## Calculating the odds ratio and 95% confidence interval for each predictor variable from the hurdle NB model's results


```{r}
exp(cbind(Odds_Ratio = coef(hurdlenb.model_all10), confint(hurdlenb.model_all10)))
```

```{r}
tbl_regression(hurdlenb.model_all10, exponentiate = TRUE)
```




```{r}
confint(zinb.model_all10, level=0.95)
```


## The likelihood ratio test for nested models

```{r}
lrtest(poisson.model_all10, poisson.model_all5)
```

```{r}
lrtest(negbin.model_all10, negbin.model_all5)
```

```{r}
lrtest(zip.model_all10, zip.model_all5)
```

```{r}
lrtest(zinb.model_all10, zinb.model_all5)
```

```{r}
lrtest(hurdlep.model_all10, hurdlep.model_all5)
```

```{r}
lrtest(hurdlenb.model_all10,hurdlenb.model_all5)
```

## The vuoung test between ZINB and hurdle NB (ZINB & HNB)

```{r}
vuong(hurdlenb.model_all10, zinb.model_all10)
```

```{r}
vuong(hurdlenb.model_all10, hurdlep.model_all10)
```

```{r}
lrtest(zinb.model_all10)
```

```{r}
lrtest(poisson.model_all10)
```


```{r}
logistic.display(poisson.model_all10)
```
```{r}
exp(cbind(Exponentiated_Odds_Ratio=coef(zinb.model_low10), confint(zinb.model_low10)))
```




